---
title: "Final Project PS239T: Polarity"
author: "Biz Herman"
date: "December 9, 2015"
output: html_document
---

# First we have to prepare the data.

```{r, cache=TRUE}

# set working directory
setwd("/Users/elizabeth/Documents/Berkeley/PS239T/ps239T-final-project")

# import the csv with the articles
PP_NYT <- read.csv("/Users/elizabeth/Documents/Berkeley/PS239T/ps239T-final-project/Data/PP_NYT.csv")

# subset the variables that are meaningful, create new dataset to work with
pp <- subset(PP_NYT, select=c(DATE, TITLE, LENGTH, GRAPHIC, SECTION, BYLINE, DATELINE, TEXT))
names(pp)

# create a new date variable that will represent the article date in YYYY-MM-DD format
date <- as.character(pp$DATE)
betterDates <- as.Date(date, format = "%B %d, %Y")
pp$date.num <- betterDates

# create a variable for just the year each article was published
year = NULL
pp$year <- substr(pp$date.num, 1, 4)

# create a variable for the year and month each article was published
yearmonth = NULL
pp$yearmonth <- substr(pp$date.num, 1, 7)

```

# Polarity Analysis

Now we're going to look at the polarity of the articles on Planned Parenthood, as it changes over time. Though we're using this analysis on Planned Parenthood articles, we could really use this on any corpus, to analyze how the polarity of a set of documents have changed over a given variable---be it time, from document to document, by person, and so on.

```{r, cache = TRUE, results="hide"}

# load the libraries we will need for this section
library(qdap) # quantiative discourse analysis of transcripts
library(data.table) # for easier data manipulation
library(scales) # to help us plot
library(plyr)

# subset the data for 2009 and later, for an initial analysis of the polarity trends
pp.2010 <- subset(pp, pp$year > 2009)

# then run this line of code below (i've had to hide the entire chunk below, because it was printing the entire output)
# split the texts into individual sentences
# pp.split <- sentSplit(pp.2010, "TEXT")

```


```{r, cache = TRUE, results="hide", include=FALSE}

# split the texts into individual sentences
pp.split <- sentSplit(pp.2010, "TEXT")

```


```{r, cache=TRUE, warning=FALSE}

pp.split$yearmonth <- substr(pp.split$date.num, 1, 7)

# look at the result
head(truncdf(pp.split),10)

# calculate polarity / sentiment of each line
# (ppdat <- with(pp.split, polarity(TEXT, TITLE)))

ppdat.2 <- with(pp.split, polarity(TEXT, yearmonth))
head(ppdat.2$group$ave.polarity)

# look at the results
counts(ppdat.2)[1:10,]

# 

```

We can represent this relationship graphically, as follows:

```{r}

library(ggplot2) # plotting discourse data

# plot it
pp.plot <- data.frame(yearmonth = ppdat.2$group$yearmonth, pol = ppdat.2$group$ave.polarity)

colorado <- subset(pp.plot, yearmonth == "2015-11")
lilarose <- subset(pp.plot, yearmonth == "2011-02")
cmp <- subset(pp.plot, yearmonth == "2015-08")

ggplot(data = pp.plot, aes(x = yearmonth, y = pol)) + geom_point() + geom_smooth(aes(group=1)) +
  theme(axis.text.x = element_text(angle = -90)) + 
  ylab("Average Polarity") + xlab("Date") + 
  geom_point(data=colorado, colour="red") +
  geom_text(data=colorado, label="CO PP Shooting", vjust=1.5, hjust = 1, color = "red") + 
  geom_point(data=lilarose, colour="blue") +
  geom_text(data=lilarose, label="LiveAction Sex Traffic", vjust=1.5, hjust = 1, color = "blue") +
  geom_point(data=cmp, colour="purple") +
  geom_text(data=cmp, label="CMP Release", vjust=1.5, hjust = 1, color = "purple") +
  geom_hline(aes()) + 
  geom_vline(xintercept=25, linetype="dashed", color = "blue") + 
  geom_vline(xintercept=79, linetype="dashed", color = "purple") + 
  geom_vline(xintercept=82, linetype="dashed", color = "red") +
  scale_x_discrete(breaks = levels(ppdat.2$group$yearmonth)[c(T, rep(F, 2))])

```

Let's expand the timeline to the beginning of 2000.

```{r, cache = TRUE, }

# subset the data for 2009 and later, for an initial analysis of the polarity trends
pp.2000 <- subset(pp, pp$year > 1999)

# then run this line of code below (i've had to hide the entire chunk below, because it was printing the entire output)
# split the texts into individual sentences
# pp.split.2000 <- sentSplit(pp.2000, "TEXT")

```


```{r, cache = TRUE, results="hide", include=FALSE}

# split the texts into individual sentences
pp.split.2000 <- sentSplit(pp.2000, "TEXT")

```



```{r, cache = TRUE, warning=FALSE}

pp.split.2000$yearmonth <- substr(pp.split.2000$date.num, 1, 7)

# look at the result
head(truncdf(pp.split.2000),10)

# calculate polarity / sentiment of each line
# (ppdat <- with(pp.split, polarity(TEXT, TITLE)))

ppdat.2000 <- with(pp.split.2000, polarity(TEXT, yearmonth))
head(ppdat.2000$group$ave.polarity)

# look at the results
counts(ppdat.2)[1:10,]

# plot it
pp.plot.2000 <- data.frame(yearmonth = ppdat.2000$group$yearmonth, pol = ppdat.2000$group$ave.polarity)

colorado <- subset(pp.plot.2000, yearmonth == "2015-11")
lilarose <- subset(pp.plot.2000, yearmonth == "2011-02")
cmp <- subset(pp.plot.2000, yearmonth == "2015-08")

ggplot(data = pp.plot.2000, aes(x = yearmonth, y = pol)) + geom_point() + geom_smooth(aes(group=1)) +
  theme(axis.text.x = element_text(angle = -90)) + 
  ylab("Average Polarity") + xlab("Date") + 
  geom_point(data=colorado, colour="red") +
  geom_text(data=colorado, label="CO PP Shooting", vjust=1.5, hjust = 1, color = "red") + 
  geom_point(data=lilarose, colour="blue") +
  geom_text(data=lilarose, label="LiveAction Sex Traffic", vjust=1.5, hjust = 1, color = "blue") +
  geom_point(data=cmp, colour="purple") +
  geom_text(data=cmp, label="CMP Release", vjust=1.5, hjust = 1, color = "purple") +
  geom_hline(aes()) + 
  geom_vline(xintercept=133, linetype="dashed", color = "blue") + 
  geom_vline(xintercept=187, linetype="dashed", color = "purple") + 
  geom_vline(xintercept=190, linetype="dashed", color = "red") +
  scale_x_discrete(breaks = levels(ppdat.2000$group$yearmonth)[c(T, rep(F, 4))])
  
```


Now we can complete this analysis on a larger subset of the data, from 1982 to the present.

```{r, cache=TRUE, results="hide"}

# subset the data for 1982 and later, for an initial analysis of the polarity trends
pp.1982 <- subset(pp, pp$year > 1982)

# split the texts into individual sentences
# cannot suppress this output for some reason!! to get beyond the immense amount of output, command-F "pp.split.1982$yearmonth <- substr(pp.split.1982$date.num, 1, 7)", which is the next line of code!

# then run this line of code below (i've had to hide the entire chunk below, because it was printing the entire output)
# split the texts into individual sentences
# pp.split.1982 <- sentSplit(pp.1982, "TEXT")

```


```{r, cache = TRUE, results="hide", include=FALSE}

# split the texts into individual sentences
pp.split.1982 <- sentSplit(pp.1982, "TEXT")

```


```{r, cache=TRUE, warning=FALSE}

pp.split.1982$yearmonth <- substr(pp.split.1982$date.num, 1, 7)

# look at the result
head(truncdf(pp.split.1982),10)

# calculate polarity / sentiment of each line
ppdat.1982 <- with(pp.split.1982, polarity(TEXT, yearmonth))
head(ppdat.1982$group$ave.polarity)

# look at the results
counts(ppdat.1982)[1:10,]

```

And plot it.

```{r}

library(ggplot2) # plotting discourse data

# plot it
pp.plot.1982 <- data.frame(yearmonth = ppdat.1982$group$yearmonth, pol = ppdat.1982$group$ave.polarity)

ggplot(data = pp.plot.1982, aes(x = yearmonth, y = pol)) + geom_point() + geom_smooth(aes(group=1)) +
  theme(axis.text.x = element_text(angle = -90)) + 
  ylab("Average Polarity") + xlab("Date") +
  geom_hline(aes()) + 
  scale_x_discrete(breaks = levels(ppdat.1982$group$yearmonth)[c(T, rep(F, 7))])

```

