---
title: "Final Project PS239T: Topic Modeling"
author: "Biz Herman"
date: "December 9, 2015"
output: html_document
---

# First we have to prepare the data.

```{r}

# import the csv with the articles
PP_NYT <- read.csv("/Users/elizabeth/Documents/Berkeley/PS239T/ps239T-final-project/Data/PP_NYT.csv")

# subset the variables that are meaningful, create new dataset to work with
pp <- subset(PP_NYT, select=c(DATE, TITLE, LENGTH, GRAPHIC, SECTION, BYLINE, DATELINE, TEXT))
names(pp)

# create a new date variable that will represent the article date in YYYY-MM-DD format
date <- as.character(pp$DATE)
betterDates <- as.Date(date, format = "%B %d, %Y")
pp$date.num <- betterDates

# create a variable for just the year each article was published
year = NULL
pp$year <- substr(pp$date.num, 1, 4)

# create a variable for the year and month each article was published
yearmonth = NULL
pp$yearmonth <- substr(pp$date.num, 1, 7)

```

# Topic Modeling Analysis

Now we're going to look at the polarity of the articles on Planned Parenthood, as it changes over time. Though we're using this analysis on Planned Parenthood articles, we could really use this on any corpus, to analyze how the polarity of a set of documents have changed over a given variable---be it time, from document to document, by person, and so on.

```{r}

set.seed(1234)

# load the libraries we will need for this section
library(mallet) # a wrapper around the Java machine learning tool MALLET
library(wordcloud) # to visualize wordclouds

# subset the data for 2009 and later, for an initial analysis of the topic models
pp.2010 <- subset(pp, pp$year > 2009)

# we first have to create an 'id' column
pp.2010$id <- rownames(pp.2010)

# remove punctuation
pp.2010$TEXT <- gsub(pattern="[[:punct:]]",replacement=" ",pp.2010$TEXT)


# load data into mallet
mallet.instances <- mallet.import(pp.2010$id, pp.2010$TEXT, "/Users/elizabeth/Documents/Berkeley/PS239T/ps239T-final-project/Data/stoplist.csv", FALSE, token.regexp="[\\p{L}']+")

# choose the number of topics to model
n.topics = 10

# create a topic trainer object
topic.model <- MalletLDA(n.topics)

# load the documents
topic.model$loadDocuments(mallet.instances)

# get the vocabulary, and some statistics about word frequencies; after running this code once through, i went back and re-curated the stop word lists, to remove some of the more frequently used words that weren't otherwise caught
vocabulary <- topic.model$getVocabulary()
word.freqs <- mallet.word.freqs(topic.model)

# examine some of the vocabulary
word.freqs[1:50,]

# the most frequently used words
word.freqs.ordered <- word.freqs[order(-word.freqs$term.freq), ]
head(word.freqs.ordered)

# optimize hyperparameters every 20 iterations, after 50 burn-in iterations
topic.model$setAlphaOptimization(20, 50)

# now train a model, specifying the number of iterations
topic.model$train(100)

# get the probability of topics in documents and the probability of words in topics; by default the functions return word counts, so to get the probabilities we can normalize and add smoothing, in order to ensure that nothing has a probability of exactly 0
doc.topics <- mallet.doc.topics(topic.model, smoothed=T, normalized=T)
topic.words <- mallet.topic.words(topic.model, smoothed=T, normalized=T)

# what are the top words in topic 5?
mallet.top.words(topic.model, topic.words[4,])

# create a vector containing short names for the topics
topics.labels <- rep("", n.topics)
for (topic in 1:n.topics) topics.labels[topic] <- paste(mallet.top.words(topic.model, topic.words[topic,], num.top.words=5)$words, collapse=" ")

# have a look at keywords for each topic
topics.labels

# show the first few document titles with at least .25 of its content devoted to topic 1
head(pp.2010$TITLE[ doc.topics[1,] > 0.25 ],10)

# create a vector that has the title of the most representative text for each topic
topics.articles <- rep("", n.topics)
for (i in 1:n.topics) topics.articles[i] <- paste(pp.2010[which.max(doc.topics[i, ]), ]$TITLE)

# weirdly, many of the topics have the same text that is most representative
topics.articles

# now let's look at how topics differ across different years?
topic.words.2009 <- mallet.subset.topic.words(topic.model, pp.2010$year == 2009, smoothed=T, normalized=T)
topic.words.2010 <- mallet.subset.topic.words(topic.model, pp.2010$year == 2010, smoothed=T, normalized=T)
topic.words.2011 <- mallet.subset.topic.words(topic.model, pp.2010$year == 2011, smoothed=T, normalized=T)
topic.words.2012 <- mallet.subset.topic.words(topic.model, pp.2010$year == 2012, smoothed=T, normalized=T)
topic.words.2013 <- mallet.subset.topic.words(topic.model, pp.2010$year == 2013, smoothed=T, normalized=T)
topic.words.2014 <- mallet.subset.topic.words(topic.model, pp.2010$year == 2014, smoothed=T, normalized=T)
topic.words.2015 <- mallet.subset.topic.words(topic.model, pp.2010$year == 2015, smoothed=T, normalized=T)

topics.labels.2010 <- rep("", n.topics)
for (topic in 1:n.topics) topics.labels.2010[topic] <- paste(mallet.top.words(topic.model, topic.words.2010[topic,], num.top.words=5)$words, collapse=" ")
topics.labels.2010

topics.labels.2011 <- rep("", n.topics)
for (topic in 1:n.topics) topics.labels.2011[topic] <- paste(mallet.top.words(topic.model, topic.words.2011[topic,], num.top.words=5)$words, collapse=" ")
topics.labels.2011

topics.labels.2012 <- rep("", n.topics)
for (topic in 1:n.topics) topics.labels.2012[topic] <- paste(mallet.top.words(topic.model, topic.words.2012[topic,], num.top.words=5)$words, collapse=" ")
topics.labels.2012

topics.labels.2013 <- rep("", n.topics)
for (topic in 1:n.topics) topics.labels.2013[topic] <- paste(mallet.top.words(topic.model, topic.words.2013[topic,], num.top.words=5)$words, collapse=" ")
topics.labels.2013

topics.labels.2014 <- rep("", n.topics)
for (topic in 1:n.topics) topics.labels.2014[topic] <- paste(mallet.top.words(topic.model, topic.words.2014[topic,], num.top.words=5)$words, collapse=" ")
topics.labels.2014

topics.labels.2015 <- rep("", n.topics)
for (topic in 1:n.topics) topics.labels.2015[topic] <- paste(mallet.top.words(topic.model, topic.words.2015[topic,], num.top.words=5)$words, collapse=" ")
topics.labels.2015

# vectorize them
t.2010 <- as.vector(topics.labels.2010)
t.2011 <- as.vector(topics.labels.2011)
t.2012 <- as.vector(topics.labels.2012)
t.2013 <- as.vector(topics.labels.2013)
t.2014 <- as.vector(topics.labels.2014)
t.2015 <- as.vector(topics.labels.2015)

# view all the topics as they change over the years
topics.over.time <- cbind(t.2010, t.2011, t.2012, t.2013, t.2014, t.2015)

# look at each topic individually -- the first topic over the years
topics.over.time[1, ]

# the second!
topics.over.time[2, ]

# the third
topics.over.time[3, ]

# the fourth
topics.over.time[4, ]

# the fifth
topics.over.time[5, ]

# the sixth
topics.over.time[6, ]

# the seventh
topics.over.time[7, ]

# the eighth
topics.over.time[8, ]

# the ninth
topics.over.time[9, ]

# the tenth
topics.over.time[10, ]

```

We can represent this relationship visually, as follows:

```{r, warning=FALSE}

# with the wordcloud package
topic.num <- 1
num.top.words<-100
topic.top.words <- mallet.top.words(topic.model, topic.words[1,], 100)
wordcloud(topic.top.words$words, topic.top.words$weights, c(4,.8), rot.per=0, random.order=F, colors = "red", ordered.colors = T)

num.topics<-10
num.top.words<-25
for(i in 1:num.topics){
  topic.top.words <- mallet.top.words(topic.model, topic.words[i,], num.top.words)
  wordcloud(topic.top.words$words, topic.top.words$weights, c(4,.8), rot.per=0, random.order=F, colors = "red", ordered.colors = T)
}

```

And creating a cluster dendogram.

```{r}

# from http://www.cs.princeton.edu/~mimno/R/clustertrees.R

# transpose and normalize the doc topics
topic.docs <- t(doc.topics)
topic.docs <- topic.docs / rowSums(topic.docs)
write.csv(topic.docs, "/Users/elizabeth/Documents/Berkeley/PS239T/ps239T-final-project/Results/topic-docs.csv")

# Get a vector containing short names for the topics
topics.labels <- rep("", n.topics)
for (topic in 1:n.topics) topics.labels[topic] <- paste(mallet.top.words(topic.model, topic.words[topic,],
                                                                         num.top.words=5)$words, collapse=" ")
# have a look at keywords for each topic
topics.labels
write.csv(topics.labels, "/Users/elizabeth/Documents/Berkeley/PS239T/ps239T-final-project/Results/topic-labels.csv")

# create data.frame with columns as docs and rows as topics
topic_docs <- data.frame(topic.docs)
names(topic_docs) <- pp.2010$id

# cluster based on shared words
plot(hclust(dist(topic.words)), labels=topics.labels)

```

Now we can complete this analysis on a larger subset of the data, from 1982 to the present.

```{r}

set.seed(12345)

# load the libraries we will need for this section
library(mallet) # a wrapper around the Java machine learning tool MALLET
library(wordcloud) # to visualize wordclouds

# subset the data for 1982 and later, the dates for which we have the complete data
pp.1982 <- subset(pp, pp$year > 1982)

# we first have to create an 'id' column
pp.1982$id <- rownames(pp.1982)

# remove punctuation
pp.1982$TEXT <- gsub(pattern="[[:punct:]]",replacement=" ", pp.1982$TEXT)

# load data into mallet
mallet.instances <- mallet.import(pp.1982$id, pp.1982$TEXT, "/Users/elizabeth/Documents/Berkeley/PS239T/ps239T-final-project/Data/stoplist.csv", FALSE, token.regexp="[\\p{L}']+")

# decide what number of topics to model
n.topics = 10

# create a topic trainer object.
topic.model <- MalletLDA(n.topics)

# load our documents
topic.model$loadDocuments(mallet.instances)

# get the vocabulary, and some statistics about word frequencies. these may be useful in further curating the stopword list.
vocabulary <- topic.model$getVocabulary()
word.freqs <- mallet.word.freqs(topic.model)

# examine some of the vocabulary
word.freqs[1:50,]

# the most frequently used words
word.freqs.ordered <- word.freqs[order(-word.freqs$term.freq), ]
head(word.freqs.ordered)

# optimize hyperparameters every 20 iterations, after 50 burn-in iterations.
topic.model$setAlphaOptimization(20, 50)

# now train a model. Note that hyperparameter optimization is on, by default. We can specify the number of iterations. Here we'll use a large-ish round number.
topic.model$train(100)

# get the probability of topics in documents and the probability of words in topics; by default the functions return word counts, so to get the probabilities we can normalize and add smoothing, in order to ensure that nothing has a probability of exactly 0
doc.topics <- mallet.doc.topics(topic.model, smoothed=T, normalized=T)
topic.words <- mallet.topic.words(topic.model, smoothed=T, normalized=T)

# what are the top words in topic 7?
mallet.top.words(topic.model, topic.words[6,])

# Get a vector containing short names for the topics
topics.labels <- rep("", n.topics)
for (topic in 1:n.topics) topics.labels[topic] <- paste(mallet.top.words(topic.model, topic.words[topic,], num.top.words=5)$words, collapse=" ")

# have a look at keywords for each topic
topics.labels

# show the first few document titles with at least .25 of its content devoted to topic 1
head(pp.1982$TITLE[ doc.topics[1,] > 0.25 ],10)

# create a vector that has the title of the most representative text for each topic
topics.articles <- rep("", n.topics)
for (i in 1:n.topics) topics.articles[i] <- paste(pp.1982[which.max(doc.topics[i, ]), ]$TITLE)

# weirdly, many of the topics have the same text that is most representative
topics.articles

# now let's look at how topics differ across different years?
topic.words.1983 <- mallet.subset.topic.words(topic.model, pp.1982$year == 1983, smoothed=T, normalized=T)
topic.words.1984 <- mallet.subset.topic.words(topic.model, pp.1982$year == 1984, smoothed=T, normalized=T)
topic.words.1985 <- mallet.subset.topic.words(topic.model, pp.1982$year == 1985, smoothed=T, normalized=T)
topic.words.1986 <- mallet.subset.topic.words(topic.model, pp.1982$year == 1986, smoothed=T, normalized=T)
topic.words.1987 <- mallet.subset.topic.words(topic.model, pp.1982$year == 1987, smoothed=T, normalized=T)
topic.words.1988 <- mallet.subset.topic.words(topic.model, pp.1982$year == 1988, smoothed=T, normalized=T)
topic.words.1989 <- mallet.subset.topic.words(topic.model, pp.1982$year == 1989, smoothed=T, normalized=T)
topic.words.1990 <- mallet.subset.topic.words(topic.model, pp.1982$year == 1990, smoothed=T, normalized=T)
topic.words.1991 <- mallet.subset.topic.words(topic.model, pp.1982$year == 1991, smoothed=T, normalized=T)
topic.words.1992 <- mallet.subset.topic.words(topic.model, pp.1982$year == 1992, smoothed=T, normalized=T)
topic.words.1993 <- mallet.subset.topic.words(topic.model, pp.1982$year == 1993, smoothed=T, normalized=T)
topic.words.1994 <- mallet.subset.topic.words(topic.model, pp.1982$year == 1994, smoothed=T, normalized=T)
topic.words.1995 <- mallet.subset.topic.words(topic.model, pp.1982$year == 1995, smoothed=T, normalized=T)
topic.words.1996 <- mallet.subset.topic.words(topic.model, pp.1982$year == 1996, smoothed=T, normalized=T)
topic.words.1997 <- mallet.subset.topic.words(topic.model, pp.1982$year == 1997, smoothed=T, normalized=T)
topic.words.1998 <- mallet.subset.topic.words(topic.model, pp.1982$year == 1998, smoothed=T, normalized=T)
topic.words.1999 <- mallet.subset.topic.words(topic.model, pp.1982$year == 1999, smoothed=T, normalized=T)
topic.words.2000 <- mallet.subset.topic.words(topic.model, pp.1982$year == 2000, smoothed=T, normalized=T)
topic.words.2001 <- mallet.subset.topic.words(topic.model, pp.1982$year == 2001, smoothed=T, normalized=T)
topic.words.2002 <- mallet.subset.topic.words(topic.model, pp.1982$year == 2002, smoothed=T, normalized=T)
topic.words.2003 <- mallet.subset.topic.words(topic.model, pp.1982$year == 2003, smoothed=T, normalized=T)
topic.words.2004 <- mallet.subset.topic.words(topic.model, pp.1982$year == 2004, smoothed=T, normalized=T)
topic.words.2005 <- mallet.subset.topic.words(topic.model, pp.1982$year == 2005, smoothed=T, normalized=T)
topic.words.2006 <- mallet.subset.topic.words(topic.model, pp.1982$year == 2006, smoothed=T, normalized=T)
topic.words.2007 <- mallet.subset.topic.words(topic.model, pp.1982$year == 2007, smoothed=T, normalized=T)
topic.words.2008 <- mallet.subset.topic.words(topic.model, pp.1982$year == 2008, smoothed=T, normalized=T)
topic.words.2009 <- mallet.subset.topic.words(topic.model, pp.1982$year == 2009, smoothed=T, normalized=T)
topic.words.2010 <- mallet.subset.topic.words(topic.model, pp.1982$year == 2010, smoothed=T, normalized=T)
topic.words.2011 <- mallet.subset.topic.words(topic.model, pp.1982$year == 2011, smoothed=T, normalized=T)
topic.words.2012 <- mallet.subset.topic.words(topic.model, pp.1982$year == 2012, smoothed=T, normalized=T)
topic.words.2013 <- mallet.subset.topic.words(topic.model, pp.1982$year == 2013, smoothed=T, normalized=T)
topic.words.2014 <- mallet.subset.topic.words(topic.model, pp.1982$year == 2014, smoothed=T, normalized=T)
topic.words.2015 <- mallet.subset.topic.words(topic.model, pp.1982$year == 2015, smoothed=T, normalized=T)

topics.labels.1983 <- rep("", n.topics)
for (topic in 1:n.topics) topics.labels.1983[topic] <- paste(mallet.top.words(topic.model, topic.words.1983[topic,], num.top.words=5)$words, collapse=" ")

topics.labels.1984 <- rep("", n.topics)
for (topic in 1:n.topics) topics.labels.1984[topic] <- paste(mallet.top.words(topic.model, topic.words.1984[topic,], num.top.words=5)$words, collapse=" ")

topics.labels.1985 <- rep("", n.topics)
for (topic in 1:n.topics) topics.labels.1985[topic] <- paste(mallet.top.words(topic.model, topic.words.1985[topic,], num.top.words=5)$words, collapse=" ")

topics.labels.1986 <- rep("", n.topics)
for (topic in 1:n.topics) topics.labels.1986[topic] <- paste(mallet.top.words(topic.model, topic.words.1986[topic,], num.top.words=5)$words, collapse=" ")

topics.labels.1987 <- rep("", n.topics)
for (topic in 1:n.topics) topics.labels.1987[topic] <- paste(mallet.top.words(topic.model, topic.words.1987[topic,], num.top.words=5)$words, collapse=" ")

topics.labels.1988 <- rep("", n.topics)
for (topic in 1:n.topics) topics.labels.1988[topic] <- paste(mallet.top.words(topic.model, topic.words.1988[topic,], num.top.words=5)$words, collapse=" ")

topics.labels.1989 <- rep("", n.topics)
for (topic in 1:n.topics) topics.labels.1989[topic] <- paste(mallet.top.words(topic.model, topic.words.1989[topic,], num.top.words=5)$words, collapse=" ")

topics.labels.1990 <- rep("", n.topics)
for (topic in 1:n.topics) topics.labels.1990[topic] <- paste(mallet.top.words(topic.model, topic.words.1990[topic,], num.top.words=5)$words, collapse=" ")

topics.labels.1991 <- rep("", n.topics)
for (topic in 1:n.topics) topics.labels.1991[topic] <- paste(mallet.top.words(topic.model, topic.words.1991[topic,], num.top.words=5)$words, collapse=" ")

topics.labels.1992 <- rep("", n.topics)
for (topic in 1:n.topics) topics.labels.1992[topic] <- paste(mallet.top.words(topic.model, topic.words.1992[topic,], num.top.words=5)$words, collapse=" ")

topics.labels.1993 <- rep("", n.topics)
for (topic in 1:n.topics) topics.labels.1993[topic] <- paste(mallet.top.words(topic.model, topic.words.1993[topic,], num.top.words=5)$words, collapse=" ")

topics.labels.1994 <- rep("", n.topics)
for (topic in 1:n.topics) topics.labels.1994[topic] <- paste(mallet.top.words(topic.model, topic.words.1994[topic,], num.top.words=5)$words, collapse=" ")

topics.labels.1995 <- rep("", n.topics)
for (topic in 1:n.topics) topics.labels.1995[topic] <- paste(mallet.top.words(topic.model, topic.words.1995[topic,], num.top.words=5)$words, collapse=" ")

topics.labels.1996 <- rep("", n.topics)
for (topic in 1:n.topics) topics.labels.1996[topic] <- paste(mallet.top.words(topic.model, topic.words.1996[topic,], num.top.words=5)$words, collapse=" ")

topics.labels.1997 <- rep("", n.topics)
for (topic in 1:n.topics) topics.labels.1997[topic] <- paste(mallet.top.words(topic.model, topic.words.1997[topic,], num.top.words=5)$words, collapse=" ")

topics.labels.1998 <- rep("", n.topics)
for (topic in 1:n.topics) topics.labels.1998[topic] <- paste(mallet.top.words(topic.model, topic.words.1998[topic,], num.top.words=5)$words, collapse=" ")

topics.labels.1999 <- rep("", n.topics)
for (topic in 1:n.topics) topics.labels.1999[topic] <- paste(mallet.top.words(topic.model, topic.words.1999[topic,], num.top.words=5)$words, collapse=" ")

topics.labels.2000 <- rep("", n.topics)
for (topic in 1:n.topics) topics.labels.2000[topic] <- paste(mallet.top.words(topic.model, topic.words.2000[topic,], num.top.words=5)$words, collapse=" ")

topics.labels.2001 <- rep("", n.topics)
for (topic in 1:n.topics) topics.labels.2001[topic] <- paste(mallet.top.words(topic.model, topic.words.2001[topic,], num.top.words=5)$words, collapse=" ")

topics.labels.2002 <- rep("", n.topics)
for (topic in 1:n.topics) topics.labels.2002[topic] <- paste(mallet.top.words(topic.model, topic.words.2002[topic,], num.top.words=5)$words, collapse=" ")

topics.labels.2003 <- rep("", n.topics)
for (topic in 1:n.topics) topics.labels.2003[topic] <- paste(mallet.top.words(topic.model, topic.words.2003[topic,], num.top.words=5)$words, collapse=" ")

topics.labels.2004 <- rep("", n.topics)
for (topic in 1:n.topics) topics.labels.2004[topic] <- paste(mallet.top.words(topic.model, topic.words.2004[topic,], num.top.words=5)$words, collapse=" ")

topics.labels.2005 <- rep("", n.topics)
for (topic in 1:n.topics) topics.labels.2005[topic] <- paste(mallet.top.words(topic.model, topic.words.2005[topic,], num.top.words=5)$words, collapse=" ")

topics.labels.2006 <- rep("", n.topics)
for (topic in 1:n.topics) topics.labels.2006[topic] <- paste(mallet.top.words(topic.model, topic.words.2006[topic,], num.top.words=5)$words, collapse=" ")

topics.labels.2007 <- rep("", n.topics)
for (topic in 1:n.topics) topics.labels.2007[topic] <- paste(mallet.top.words(topic.model, topic.words.2007[topic,], num.top.words=5)$words, collapse=" ")

topics.labels.2008 <- rep("", n.topics)
for (topic in 1:n.topics) topics.labels.2008[topic] <- paste(mallet.top.words(topic.model, topic.words.2008[topic,], num.top.words=5)$words, collapse=" ")

topics.labels.2009 <- rep("", n.topics)
for (topic in 1:n.topics) topics.labels.2009[topic] <- paste(mallet.top.words(topic.model, topic.words.2009[topic,], num.top.words=5)$words, collapse=" ")

topics.labels.2010 <- rep("", n.topics)
for (topic in 1:n.topics) topics.labels.2010[topic] <- paste(mallet.top.words(topic.model, topic.words.2010[topic,], num.top.words=5)$words, collapse=" ")

topics.labels.2011 <- rep("", n.topics)
for (topic in 1:n.topics) topics.labels.2011[topic] <- paste(mallet.top.words(topic.model, topic.words.2011[topic,], num.top.words=5)$words, collapse=" ")

topics.labels.2012 <- rep("", n.topics)
for (topic in 1:n.topics) topics.labels.2012[topic] <- paste(mallet.top.words(topic.model, topic.words.2012[topic,], num.top.words=5)$words, collapse=" ")

topics.labels.2013 <- rep("", n.topics)
for (topic in 1:n.topics) topics.labels.2013[topic] <- paste(mallet.top.words(topic.model, topic.words.2013[topic,], num.top.words=5)$words, collapse=" ")

topics.labels.2014 <- rep("", n.topics)
for (topic in 1:n.topics) topics.labels.2014[topic] <- paste(mallet.top.words(topic.model, topic.words.2014[topic,], num.top.words=5)$words, collapse=" ")

topics.labels.2015 <- rep("", n.topics)
for (topic in 1:n.topics) topics.labels.2015[topic] <- paste(mallet.top.words(topic.model, topic.words.2015[topic,], num.top.words=5)$words, collapse=" ")

# vectorize them
t.1983 <- as.vector(topics.labels.1983)
t.1984 <- as.vector(topics.labels.1984)
t.1985 <- as.vector(topics.labels.1985)
t.1986 <- as.vector(topics.labels.1986)
t.1987 <- as.vector(topics.labels.1987)
t.1988 <- as.vector(topics.labels.1988)
t.1989 <- as.vector(topics.labels.1989)
t.1990 <- as.vector(topics.labels.1990)
t.1991 <- as.vector(topics.labels.1991)
t.1992 <- as.vector(topics.labels.1992)
t.1993 <- as.vector(topics.labels.1993)
t.1994 <- as.vector(topics.labels.1994)
t.1995 <- as.vector(topics.labels.1995)
t.1996 <- as.vector(topics.labels.1996)
t.1997 <- as.vector(topics.labels.1997)
t.1998 <- as.vector(topics.labels.1998)
t.1999 <- as.vector(topics.labels.1999)
t.2001 <- as.vector(topics.labels.2001)
t.2002 <- as.vector(topics.labels.2002)
t.2003 <- as.vector(topics.labels.2003)
t.2004 <- as.vector(topics.labels.2004)
t.2005 <- as.vector(topics.labels.2005)
t.2000 <- as.vector(topics.labels.2000)
t.2001 <- as.vector(topics.labels.2001)
t.2002 <- as.vector(topics.labels.2002)
t.2003 <- as.vector(topics.labels.2003)
t.2004 <- as.vector(topics.labels.2004)
t.2005 <- as.vector(topics.labels.2005)
t.2006 <- as.vector(topics.labels.2006)
t.2007 <- as.vector(topics.labels.2007)
t.2008 <- as.vector(topics.labels.2008)
t.2009 <- as.vector(topics.labels.2009)
t.2010 <- as.vector(topics.labels.2010)
t.2011 <- as.vector(topics.labels.2011)
t.2012 <- as.vector(topics.labels.2012)
t.2013 <- as.vector(topics.labels.2013)
t.2014 <- as.vector(topics.labels.2014)
t.2015 <- as.vector(topics.labels.2015)

# create a matrix with all the topics over time
topics.over.time <- cbind(t.1983, t.1984, t.1985, t.1986, t.1987, t.1988, t.1989, t.1990, t.1991, t.1992, t.1993, t.1994, t.1995, t.1996, t.1997, t.1998, t.1999, t.2000, t.2001, t.2002, t.2003, t.2004, t.2005, t.2006, t.2007, t.2008, t.2009, t.2010, t.2011, t.2012, t.2013, t.2014, t.2015)

```

Now we can look at how the topics have changed over the years, to see if Planned Parenthood has become a more politicized issue over time, or perhaps during certain election cycles.

```{r}

# look at each topic individually -- the first topic over the years
topics.over.time[1, ]

# the second!
topics.over.time[2, ]

# the third
topics.over.time[3, ]

# the fourth
topics.over.time[4, ]

# the fifth
topics.over.time[5, ]

# the sixth
topics.over.time[6, ]

# the seventh
topics.over.time[7, ]

# the eighth
topics.over.time[8, ]

# the ninth
topics.over.time[9, ]

# the tenth
topics.over.time[10, ]

```

The ninth topic is especially interesting---it appears to track scandals or politicized issues that Planned Parenthood is embroiled in. To take a peek at how that has changed over the years, we can see that in 1985, the most common words in the topic were "bours public office called investigation." In 1993, they were "death suicide public told office." In 1999, they were "kopp smith web death site." In 2006, they were "death kline group found called." In 2012, they were "told video case web kimbrough." And in 2015, they were "tissue fetal video planned people." We can also trace changes in the topic that's about the Supreme Court, in the topic that's about election, in the topic that's about sex education---these prove to be very informative topics from which we can build interesting additional research questions!

We can also represent this topics visually, as follows:

```{r, warning=FALSE}

# with the wordcloud package
topic.num <- 1
num.top.words<-100
topic.top.words <- mallet.top.words(topic.model, topic.words[1,], 100)
wordcloud(topic.top.words$words, topic.top.words$weights, c(4,.8), rot.per=0, random.order=F, colors = "red", ordered.colors = T)

num.topics<-10
num.top.words<-25

for(i in 1:num.topics){
  topic.top.words <- mallet.top.words(topic.model, topic.words[i,], num.top.words)
  wordcloud(topic.top.words$words, topic.top.words$weights, c(4,.8), rot.per=0, random.order=F, colors = "red", ordered.colors = T)
}

```

And create a cluster dendogram.

```{r}

# from http://www.cs.princeton.edu/~mimno/R/clustertrees.R

# transpose and normalize the doc topics
topic.docs <- t(doc.topics)
topic.docs <- topic.docs / rowSums(topic.docs)
write.csv(topic.docs, "/Users/elizabeth/Documents/Berkeley/PS239T/ps239T-final-project/Results/topic-docs2.csv")

# Get a vector containing short names for the topics
topics.labels <- rep("", n.topics)
for (topic in 1:n.topics) topics.labels[topic] <- paste(mallet.top.words(topic.model, topic.words[topic,],
                                                                         num.top.words=5)$words, collapse=" ")
# have a look at keywords for each topic
topics.labels
write.csv(topics.labels, "/Users/elizabeth/Documents/Berkeley/PS239T/ps239T-final-project/Results/topic-labels2.csv")

# create data.frame with columns as docs and rows as topics
topic_docs <- data.frame(topic.docs)
names(topic_docs) <- pp.1982$id

# cluster based on shared words
png('Graphic_PPDendogram.png')
plot(hclust(dist(topic.words)), labels=topics.labels)
dev.off()

```
